---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
#install.packages("mice")
library(dplyr)
library(ggplot2)
require(plyr)
require(Hmisc)
```

```{r}
dat <- read.table("activity.csv",header=T,sep=",")

head(dat)
```
#steps    date   interval  
#<int>    <fctr>  <int>
#1	NA	2012-10-01	0	
#2	NA	2012-10-01	5	
#3	NA	2012-10-01	10	
#4	NA	2012-10-01	15	
#5	NA	2012-10-01	20	
#6	NA	2012-10-01	25	
#6 rows




```{r}
typeof(dat)
```
#[1] "list"

### create a new variable - dataFrame
```{r}
dt <- data.frame(dat)
head(dt)
```
#steps    date   interval  
#<int>    <fctr>  <int>
#1	NA	2012-10-01	0	
#2	NA	2012-10-01	5	
#3	NA	2012-10-01	10	
#4	NA	2012-10-01	15	
#5	NA	2012-10-01	20	
#6	NA	2012-10-01	25	
#6 rows



## What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
```{r}
steps_per_day <- aggregate(steps ~ date, data=dt, FUN=sum)
```


```{r}
head(steps_per_day)
```


    
#      date    steps
#     <fctr>   <int>
#1	2012-10-02	126		
#2	2012-10-03	11352		
#3	2012-10-04	12116		
#4	2012-10-05	13294		
#5	2012-10-06	15420		
#6	2012-10-07	11015		
#6 rows

## Make a histogram of the total number of steps taken each day
```{r}
hist(steps_per_day$steps, col="blue",xlab = "Total steps daily", 
        main = "Number of steps per day")
```





### Calculate the mean and median of the total number of steps taken per day
```{r}
median(steps_per_day$steps,na.rm = TRUE)
mean(steps_per_day$steps,na.rm = TRUE)
```
#[1] 10765
#[1] 10766.19


## What is the average daily activity pattern?
### 1. Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps_i <- aggregate(steps ~ interval, data = dt, FUN = mean, na.rm = TRUE)
str(steps_i)
```
#'data.frame':	288 obs. of  2 variables:
#$ interval: int  0 5 10 15 20 25 30 35 40 45 ...
#$ steps   : num  1.717 0.3396 0.1321 0.1509 0.0755 ...


## timeseries graph
```{r}
plot(steps_i$interval,steps_i$steps,type="l",xlab="time = 5 min",
        ylab="Average number of steps",main="Average daily activity pattern")
```



## 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_i <- steps_i[which.max(steps_i$steps),]
max_i
```
#interval   steps
#<int>      <dbl>
#104	835	206.1698		
#1 row


## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
cout_missing_values <- sum(is.na(dt$steps))
cout_missing_values
```
#[1] 2304



### 2. Fill in all of the missing values in the dataset


### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r}
dt_Imputed <- ddply(dt, "interval", mutate, imputed.value = impute(steps, mean))
#dt_Imputed <- ddply(dt$steps, mean)
```


```{r}
median(dt_Imputed$steps,na.rm = TRUE)
mean(dt_Imputed$steps,na.rm = TRUE)
```
#[1] 0
#[1] 37.3826

## alternative way

```{r}
fillITin <-  function(x){

x$steps[is.na(x$steps)] <- ave(x$steps,  
                                     FUN = function(z) 
                                    mean(z, na.rm = TRUE))[c(which(is.na(x$steps)))]
return(x)
}


dt_Inputed_2 <- fillITin(dt)
```


```{r}
median(dt_Inputed_2$steps,na.rm = TRUE)
mean(dt_Inputed_2$steps,na.rm = TRUE)
```
#[1] 0
#[1] 37.3826


### 3. Histogram of the total number of steps taken each day
```{r}
hist(dt_Inputed_2$steps, col="light blue",xlab = "Total steps daily", 
         main = "Number of steps per day with imputed values")
```

         
### 4. Calculate and report the mean total number of steps taken per day
```{r}
mean(dt_Inputed_2$steps,na.rm = TRUE)
```
#[1] 37.3826

### 5. Calculate and report the median total number of steps taken per day
```{r}
median(dt_Inputed_2$steps,na.rm = TRUE)
```
#[1] 0

## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
dt_Inputed_2$dateType <-  ifelse(as.POSIXlt(dt_Inputed_2$date)$wday %in% c(0,6), 'weekend', 'weekday')
```


### 2. Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
av_dt_Inputed_2 <- aggregate(steps ~ interval + dateType, data=dt_Inputed_2, mean)
ggplot(av_dt_Inputed_2, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
